---
title: "Overview of cjustin.utils"
author: "Justin Chu"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of cjustin.utils}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

cjustin.utils
-------

This package contains custom functions designed to make common tasks easier. Here I will outline example and usage of these custom functions.

Make fake dataset for illustrative purposes (derived from [here](http://svitsrv25.epfl.ch/R-doc/library/gdata/html/reorder.html)):
```{r}
library(cjustin.utils)
testDat <- data.frame( treatment = sample( c("PLACEBO", "300 MG", "600 MG", "1200 MG"),
                       100, replace=TRUE ))
knitr::kable(tail(testDat, 10), format = "markdown")
```

## writeTSV
This function is a wrapper around `write.table` that will output a tab seperated file (tsv).

```{r}
writeTSV(testDat, "testDat.tsv")

#does file exist
file.exists("testDat.tsv")
testDat2 <- read.table("testDat.tsv", sep = "\t", quote = "")

#does file load correctly
knitr::kable(tail(testDat2, 10), format = "markdown")
```

## readTSV
This function is a wrapper around `write.read` that will take a tab seperated file (tsv), a sort of compation to `readTSV`

```{r}
testDat3 <- readTSV("testDat.tsv")

#does file load correctly
knitr::kable(tail(testDat3, 10), format = "markdown")

#check if identical to original
identical(testDat3, testDat)

#cleanup
file.remove("testDat.tsv")
```

## peek
This function is like `head` or `tail` except it looks in the middle of the file rather than the start or end.

```{r}
knitr::kable(peek(testDat), format = "markdown")
```

Like `head` or `tail` you can specify how many rows you want to look at using `n` (also the second parameters):
```{r}
knitr::kable(peek(testDat, 10), format = "markdown")
```

You can also set where in the file you want to look using `loc` (value between 0 and 1, 0.5 means you start looking directly in the middle of the file):
```{r}
knitr::kable(peek(testDat, 10, 0.1), format = "markdown")
```

## formatFig
This function formats ggplot2

```{r, fig.show='hold'}
library(ggplot2)
p <- ggplot(testDat, aes(treatment)) + geom_bar()
#without formatting
p
#with formatting
p + formatFig("bioinformatics")
```

I know the figure doesn't quite look right or better than the defaults, but the point of this function was to create a theme suitable for publication (black and white, high resolution, with large text). When used for the resolution that I used these defaults worked really well.

## reorderdesc
This function uses `reorder` to change the order of the factors except reorders them in reverse order using `desc` from `dplyr`.

```{r, fig.show='hold'}
library(ggplot2)
#reorder using reorder + order
reorderedDat <- testDat
reorderedDat$treatment <- reorder(testDat$treatment, order(testDat$treatment))
reorderedDescDat <- testDat
reorderedDescDat$treatment <- reorderdesc(testDat$treatment, order(testDat$treatment))
#factors ordered in ascending order
ggplot(reorderedDat, aes(treatment)) + geom_bar()
#factors ordered in desceding order
ggplot(reorderedDescDat, aes(treatment)) + geom_bar()
```

You can see the order of the factors is reversed in my function compared to the normal `reorder`.

Reflections:
------
More than our assignment on automation, making this package reminds me of a XKCD comic related to [automation](http://xkcd.com/1319/). Each function serves as something I have to maintain as small units that are essentially functions I automated for myself. Basically I don't really know if between all the tests and documentation if I am really saving time. Negativity aside, I actually think that I will likely benefit from this task in the long run, but it is a strange thing trying to predict what functions will be useful in the future.

For each function the general process of: coding -> documentation -> updating requirement -> writing tests -> updating vignettes -> checking - although tedious, could easily become second nature, and is a good way to work since it empasizes coding with minimal redundancy and robustness (through testing).